---
title: "Predicting Wildfire Risk in Northern California"
author: "Leah Shapiro and Palak Agarwal"
date: "12/13/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
# Introduction
- explain the problem
- visualize California fires

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
```

- explain our use case

# Data
-describe the data used:
-fire perimeter
- WUI
- FVEG
- Elevation
- Slope
- Land Cover
- County
- Weather (temp, precip, humidity, wind speed)
- describe feature engineering

```{r data, message=FALSE, warning=FALSE}

```


## Exploratory Analysis
- bar plots
- correlation matrix
- maybe some maps?

```{r exploratory, message=FALSE, warning=FALSE}

```

## Describe the Spatial Process/ Space Time Process

# Model
- explain rare event problem
- list features in model
- discuss statistical significance

```{r model, message=FALSE, warning=FALSE}

```


## Goodness of Fit
-McFadden
-Density of probabilities plot
-selecting a threshold
-Confusion Matrix
-ROC Curve
-discuss accuracy

```{r fit, message=FALSE, warning=FALSE}

```

## Cross Validation
- discuss generalizability
```{r cv, message=FALSE, warning=FALSE}

```

### Spatial Cross Validation
-discuss generalizability
```{r spatial_cv, message=FALSE, warning=FALSE}

```

# Cost Benefit Analysis
```{r cba, message=FALSE, warning=FALSE}

```

# Discussion
-critique strengths and weaknesses of model

# Conclusion