---
title: "Predicting Wildfire Risk in Northern California"
author: "Leah Shapiro and Palak Agarwal"
date: "12/13/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
# Introduction
- explain the problem
- visualize California fires
-loading libraries, themes, color palettes, and functions that we will use in our analysis

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)

# LIBRARIES
library(rjson)
library(tidycensus)
library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)
library(jtools)  
library(viridis)
library(kableExtra)
library(rlist)
library(dplyr)
library(osmdata)
library(geosphere)
library(fastDummies)
library(FNN)
library(viridis)
library(stargazer)
library(pscl)
library(pROC)
library(plotROC)
library(RANN)
library(riem)
options(scipen=999)
options(tigris_class = "sf")

# THEMES AND FUNCTIONS
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}

# PALETTE
palette2 <- c("#F96167","#FCE77D")

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

q5 <- function(variable) {as.factor(ntile(variable, 5))}

# FUNCTIONS
nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <- as.matrix(measureFrom)
  measureTo_Matrix <- as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}

# READ IN DATA
all_counties <- st_read("C:/Users/owner160829a/Desktop/Graduate School/Penn/Courses/Fall 20/MUSA 508/Final Project/Geoprocessing/counties.shp") %>%
  st_transform('EPSG:2225')

fire_pt <- st_read("https://services1.arcgis.com/jUJYIo9tSA7EHvfZ/arcgis/rest/services/California_Fire_Perimeters/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson")%>%
  st_transform('EPSG:2225')

selected_counties <- st_read("https://opendata.arcgis.com/datasets/a61c138d0a6946da8d1ebb8d1c9db13a_0.geojson") %>%
  filter(COUNTY_NAME == 'Del Norte' | COUNTY_NAME == 'Siskiyou' | COUNTY_NAME == 'Humboldt' | COUNTY_NAME == 'Trinity' |
           COUNTY_NAME == 'Shasta' | COUNTY_NAME == 'Tehama' | COUNTY_NAME == 'Mendocino' | COUNTY_NAME == 'Glenn' |
           COUNTY_NAME == 'Lake' | COUNTY_NAME == 'Colusa' | COUNTY_NAME == 'Sonoma' |COUNTY_NAME == 'Napa' | COUNTY_NAME == 'Yolo') %>%
  st_transform('EPSG:2225')

```

```{r maps, message=FALSE, warning=FALSE}
fire_perimeter1019 <- fire_pt %>% filter(YEAR_=="2010"|YEAR_=="2011"|YEAR_=="2012"|YEAR_=="2013"|YEAR_=="2014"|YEAR_=="2015"|YEAR_=="2016"|YEAR_=="2017"|YEAR_=="2018"|YEAR_=="2019")

ggplot() +
  geom_sf(data = fire_perimeter1019, fill="orange", color="transparent")+
 geom_sf(data=all_counties, fill="transparent")+ 
  labs(title="California Wildfires",
       subtitle="Years 2010-2019")+
  mapTheme()

## Showing our selected counties
ggplot() +
  geom_sf(data = fire_perimeter1019, fill="orange", color="transparent")+
  geom_sf(data=all_counties, fill="transparent", size=.5, color="grey")+ 
  geom_sf(data=selected_counties, fill="transparent", color="#F96167", size=1)+
  labs(title="California Wildfires",
       subtitle="Years 2010-2019; Selected Counties in Red")+
  mapTheme()
```

-explain our use case

# Data
-describe the data used:
- fire perimeter
- WUI
- FVEG
- Elevation
- Slope
- Land Cover
- County
- Weather (temp, precip, humidity, wind speed)
- describe process in ArcMap
- explain fishnet size

```{r data, message=FALSE, warning=FALSE, cache=TRUE}
## READ IN DATA

fire_suppression_facilities <- st_read("C:/Users/owner160829a/Desktop/Graduate School/Penn/Courses/Fall 20/MUSA 508/Final Project/Geoprocessing/fire_suppression_facilities.shp")

fishnet_unclipped <- st_read("C:/Users/owner160829a/Desktop/Graduate School/Penn/Courses/Fall 20/MUSA 508/Final Project/Geoprocessing/Fishnet with Joined Data/fishnet_halfmile_joins.shp") %>%
  st_transform('EPSG:2225')

fishnet_clipped <- st_intersection(fishnet_unclipped,selected_counties)

fishnet_clipped <- fishnet_clipped %>% dplyr::select(WUI_MAJORI,FVEG_MAJOR,ELEVATION_,SLOPE_MEAN,COVER_MAJ, COUNTY_NAME, geometry) %>%
  rename (WUI_MAJ=WUI_MAJORI, 
          FVEG_MAJ=FVEG_MAJOR,
          ELEVATION_AV=ELEVATION_)

# Adding Unique IDs for each cell
fishnet_clipped$ID <-  seq.int(nrow(fishnet_clipped))
```


```{r fire_data, message=FALSE, warning=FALSE}
## Joining fire data to fishnets 
###2014-18
fire_perimeter1418 <-
  fire_pt %>%
  filter(YEAR_  == '2014' | YEAR_  == '2015' | YEAR_  == '2016' | YEAR_ =='2017' | YEAR_  == '2018') %>%
  st_transform('EPSG:2225')

clip1418 <- 
  st_intersection(st_make_valid(fire_perimeter1418),st_make_valid(fishnet_clipped)) %>%
  select(ID) %>%
  st_drop_geometry() %>%
  mutate(Fire1418 = 1) %>%
  distinct()

fishnet_clipped <-
  fishnet_clipped %>%
  left_join(., clip1418, on= 'ID') 

fishnet_clipped$Fire1418 <- ifelse(is.na(fishnet_clipped$Fire1418),0, fishnet_clipped$Fire1418)

###2019
fire_perimeter19 <-
  fire_pt %>%
  filter(YEAR_ =='2019') %>%
  st_transform('EPSG:2225')

clip19 <- 
  st_intersection(st_make_valid(fire_perimeter19),st_make_valid(fishnet_clipped)) %>%
  select(ID) %>%
  st_drop_geometry() %>%
  mutate(Fire19 = 1) %>%
  distinct()

fishnet_clipped <-
  fishnet_clipped %>%
  left_join(., clip19, on= 'ID') 

fishnet_clipped$Fire19 <- ifelse(is.na(fishnet_clipped$Fire19),0, fishnet_clipped$Fire19)
```


```{r weather_data, message=FALSE, warning=FALSE}
## WEATHER DATA

# vector 1 - of southern california station ids
weather_station_ids <- c("SIY", "CEC", "MHS", "O86", "ACV", "FOT", "RDD", "RBL", "CIC", "OVE",
                         "UKI", "MYV", "STS", "O69", "DVO", "APC", "SUU", "VCB", "EDU", "SMF", "LHM", "MYV")

# df - stations with lat/lon and name info (in addition to ids)
asos_socal_stations <- riem_stations("CA_ASOS") %>% filter(str_detect(id, paste(weather_station_ids, collapse="|")))
asos_socal_stations$weather_station_id <- asos_socal_stations$id
asos_socal_stations <-  st_as_sf(asos_socal_stations, coords = c("lon","lat"), crs = 4326, agr = "constant") %>% st_transform('EPSG:2225')
asos_socal_stations$weather_ID <-  seq.int(nrow(asos_socal_stations))

## Finding closest station
weather_coords <- 
  asos_socal_stations %>%
  select(geometry)

fishnet_coords <- 
  fishnet_clipped %>%
  select(geometry)

closest_weather_station_to_fishnet <- nn2(weather_coords, fishnet_coords, k = 1)$nn.idx

fishnet_clipped$weather_ID <- closest_weather_station_to_fishnet

ggplot()+
  geom_sf(data = selected_counties)+
  geom_sf(data = weather_coords)

# function and loop -- tried it with the vector of ids and the df of all info, but neither worked. This version below uses just the vector of station ids
get_weather_features_by_station <- function(weather_station_ids, start_year, end_year){
  
  year_vec <- seq(start_year, end_year)
  i <- 1
  weather_data_list <- list()
  for(station_id in weather_station_ids){
    print(paste("Processing station", station_id))
    for(year in year_vec){
      start_date = paste0(year, "07-01")
      end_date = paste0(year, "10-31")
      weather_data <- riem_measures(station = station_id, date_start = start_date, date_end = end_date) %>% 
        dplyr::summarise(weather_station_id = station_id,
                         year = year,
                         Max_Temp = max(tmpf, na.rm = TRUE),
                         Mean_Temp = mean(tmpf, na.rm = TRUE),
                         Mean_Precipitation = mean(p01i, na.rm = TRUE),
                         Mean_Humidity = mean(relh, na.rm = TRUE),
                         Mean_Wind_Speed = mean(sknt, na.rm = TRUE),
        ) 
      weather_data_list[[i]] <- weather_data
      i <- i + 1
    }
  }
  
  do.call("rbind", weather_data_list) 
}

weather_data2014 <- get_weather_features_by_station(weather_station_ids, 2014, 2014) %>%
  rename(Max_Temp14 = Max_Temp,
         Mean_Temp14 = Mean_Temp,
         Mean_Precipitation14 = Mean_Precipitation,
         Mean_Humidity14 = Mean_Humidity,
         Mean_Wind_Speed14 = Mean_Wind_Speed)

weather_2014 <- left_join(weather_data2014, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry) %>%
  distinct() 

fishnet_clipped <- left_join(fishnet_clipped, weather_2014, on = "weather_ID")

weather_data2015 <- get_weather_features_by_station(weather_station_ids, 2015, 2015) %>%
  rename(Max_Temp15 = Max_Temp,
         Mean_Temp15 = Mean_Temp,
         Mean_Precipitation15 = Mean_Precipitation,
         Mean_Humidity15 = Mean_Humidity,
         Mean_Wind_Speed15 = Mean_Wind_Speed)

weather_2015 <- left_join(weather_data2015, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry) %>%
  distinct()

fishnet_clipped <- left_join(fishnet_clipped, weather_2015, on = "weather_ID")

weather_data2016 <- get_weather_features_by_station(weather_station_ids, 2016, 2016) %>%
  rename(Max_Temp16 = Max_Temp,
         Mean_Temp16 = Mean_Temp,
         Mean_Precipitation16 = Mean_Precipitation,
         Mean_Humidity16 = Mean_Humidity,
         Mean_Wind_Speed16 = Mean_Wind_Speed)

weather_2016 <- left_join(weather_data2016, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry)%>%
  distinct() 

fishnet_clipped <- left_join(fishnet_clipped, weather_2016, on = "weather_ID")

weather_data2017 <- get_weather_features_by_station(weather_station_ids, 2017, 2017) %>%
  rename(Max_Temp17 = Max_Temp,
         Mean_Temp17 = Mean_Temp,
         Mean_Precipitation17 = Mean_Precipitation,
         Mean_Humidity17 = Mean_Humidity,
         Mean_Wind_Speed17 = Mean_Wind_Speed)

weather_2017 <- left_join(weather_data2017, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry)%>%
  distinct() 

fishnet_clipped <- left_join(fishnet_clipped, weather_2017, on = "weather_ID")

weather_data2018 <- get_weather_features_by_station(weather_station_ids, 2018, 2018) %>%
  rename(Max_Temp18 = Max_Temp,
         Mean_Temp18 = Mean_Temp,
         Mean_Precipitation18 = Mean_Precipitation,
         Mean_Humidity18 = Mean_Humidity,
         Mean_Wind_Speed18 = Mean_Wind_Speed)

weather_2018 <- left_join(weather_data2018, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry)%>%
  distinct() 

fishnet_clipped <- left_join(fishnet_clipped, weather_2018, on = "weather_ID")

weather_data2019 <- get_weather_features_by_station(weather_station_ids, 2019, 2019) %>%
  rename(Max_Temp19 = Max_Temp,
         Mean_Temp19 = Mean_Temp,
         Mean_Precipitation19 = Mean_Precipitation,
         Mean_Humidity19 = Mean_Humidity,
         Mean_Wind_Speed19 = Mean_Wind_Speed)

weather_2019 <- left_join(weather_data2019, asos_socal_stations, on = 'weather_station_id') %>%
  select (-weather_station_id, -id, -name, -year, -geometry)%>%
  distinct() 

fishnet_clipped <- left_join(fishnet_clipped, weather_2019, on = "weather_ID")
```

- describe feature engineering

```{r features, message=FALSE, warning=FALSE}
# FEATURE ENGINEERING
# Finding Means of Weather data
fishnet_clipped$MeanTemp1418 = (fishnet_clipped$Mean_Temp14+
                                  fishnet_clipped$Mean_Temp15+
                                  fishnet_clipped$Mean_Temp16+
                                  fishnet_clipped$Mean_Temp17+
                                  fishnet_clipped$Mean_Temp18)/5

fishnet_clipped$MeanHumidity1418 = (fishnet_clipped$Mean_Humidity14+
                                      fishnet_clipped$Mean_Humidity15+
                                      fishnet_clipped$Mean_Humidity16+
                                      fishnet_clipped$Mean_Humidity17+
                                      fishnet_clipped$Mean_Humidity18)/5

fishnet_clipped$MeanPrecip1418 = (fishnet_clipped$Mean_Precipitation14+
                                    fishnet_clipped$Mean_Precipitation15+
                                    fishnet_clipped$Mean_Precipitation16+
                                    fishnet_clipped$Mean_Precipitation17+
                                    fishnet_clipped$Mean_Precipitation18)/5

fishnet_clipped$MeanWindSpeed1418 = (fishnet_clipped$Mean_Wind_Speed14+
                                       fishnet_clipped$Mean_Wind_Speed15+
                                       fishnet_clipped$Mean_Wind_Speed16+
                                       fishnet_clipped$Mean_Wind_Speed17+
                                       fishnet_clipped$Mean_Wind_Speed18)/5

fishnet_clipped$MeanMaxTemp = (fishnet_clipped$Max_Temp14+
                                 fishnet_clipped$Max_Temp15+
                                 fishnet_clipped$Max_Temp16+
                                 fishnet_clipped$Max_Temp17+
                                 fishnet_clipped$Max_Temp18)/5

# Remove Unnecessary Variables
fishnet_clipped <- fishnet_clipped %>% dplyr::select (                                                -Mean_Temp14,-Mean_Temp15,-Mean_Temp16,-Mean_Temp17,                                                -Mean_Temp18,-Mean_Humidity14,-Mean_Humidity15,
                                                      -Mean_Humidity16,-Mean_Humidity17,-Mean_Humidity18,
                                                      -Mean_Precipitation14,-Mean_Precipitation15,                                   -Mean_Precipitation16, -Mean_Precipitation17,
                                                      -Mean_Precipitation18,-Mean_Wind_Speed14,
                                                      -Mean_Wind_Speed15,-Mean_Wind_Speed16,
                                                      -Mean_Wind_Speed17,-Mean_Wind_Speed18,
                                                      -Max_Temp14,-Max_Temp15,-Max_Temp16,-Max_Temp17,-Max_Temp18)

## Changing Integers to Characters

fishnet_clipped$WUI_MAJ <- as.factor(fishnet_clipped$WUI_MAJ)

fishnet_clipped$FVEG_MAJ <- as.factor(fishnet_clipped$FVEG_MAJ)

fishnet_clipped$COVER_MAJ <- as.factor(fishnet_clipped$COVER_MAJ)

## Fire in last 3 years
fire_perimeter1013 <-
  fire_pt %>%
  filter(YEAR_  == '2010' | YEAR_ =='2011'| YEAR_ =='2012'| YEAR_ =='2013') %>%
  st_transform('EPSG:2225')

clip1013 <- 
  st_intersection(st_make_valid(fire_perimeter1013),st_make_valid(fishnet_clipped)) %>%
  select(ID) %>%
  st_drop_geometry() %>%
  mutate(Fire1013 = 1) %>%
  distinct()

fishnet_clipped <-
  fishnet_clipped %>%
  left_join(., clip1013, on= 'ID') 

fishnet_clipped$Fire1013 <- ifelse(is.na(fishnet_clipped$Fire1013),0, fishnet_clipped$Fire1013)

## Historical fire
##intersections of fire perimeters with each fishnet cell.
fishnet_clipped <- 
  fishnet_clipped %>% 
  mutate(n_fires_intersections = lengths(st_intersects(st_make_valid(fishnet_clipped), st_make_valid(fire_pt))))

## adding a column for y/n for historical fire presence
#fishnet_clipped <-
 # fishnet_clipped %>%
  #mutate(prev_fire = ifelse(fishnet_clipped$n_fires_intersections > 0, "1", "0"))

## Categorical Features
fishnet_clipped <- fishnet_clipped %>% mutate(CoverCat = case_when(fishnet_clipped$COVER_MAJ=="1"|fishnet_clipped$COVER_MAJ=="2"|fishnet_clipped$COVER_MAJ=="4" ~ "forest",
                                              fishnet_clipped$COVER_MAJ=="6"|fishnet_clipped$COVER_MAJ=="7" ~ "shrubland",
                                              fishnet_clipped$COVER_MAJ=="8"|fishnet_clipped$COVER_MAJ=="9"|fishnet_clipped$COVER_MAJ=="10"~ "savanna_grassland",
                                              fishnet_clipped$COVER_MAJ=="11"|fishnet_clipped$COVER_MAJ=="15"|fishnet_clipped$COVER_MAJ=="17"~ "wet",
                                              fishnet_clipped$COVER_MAJ=="14"|fishnet_clipped$COVER_MAJ=="12" ~ "cropland",
                                              fishnet_clipped$COVER_MAJ=="13" ~ "urban",
                                              fishnet_clipped$COVER_MAJ=="17" ~ "barren"))


fishnet_clipped <- fishnet_clipped %>% mutate(SlopeCat = case_when(fishnet_clipped$SLOPE_MEAN < 5 ~ "low",
                                                                   fishnet_clipped$SLOPE_MEAN >=5|fishnet_clipped$SLOPE_MEAN <15 ~ "medium",
                                                                   fishnet_clipped$SLOPE_MEAN >=15 ~ "high" ))

## Replacing Land Cover type 15 (perm snow and ice) with type 17 (waterbodies)
fishnet_clipped$COVER_MAJ <- ifelse(fishnet_clipped$COVER_MAJ=="15","17",fishnet_clipped$COVER_MAJ)

## Dummy Feature
fishnet_clipped <- fishnet_clipped %>% mutate (ElevationBi = if_else(fishnet_clipped$ELEVATION_AV>3000,1,0))

## Nearest Neighbor Features
conifer_points <- fishnet_clipped %>% filter(FVEG_MAJ=="1") %>% st_centroid()

shrub_points<- fishnet_clipped %>% filter(FVEG_MAJ=="2") %>% st_centroid()

hardwood_points <- fishnet_clipped %>% filter(FVEG_MAJ=="6") %>% st_centroid()

wui_points <- fishnet_clipped %>% filter(WUI_MAJ=="4") %>% st_centroid()

fire1013_points <- fishnet_clipped %>% filter(Fire1013=="1") %>% st_centroid()

fishnet_clipped <- fishnet_clipped %>%
  mutate(
    Conifer.nn =
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(conifer_points),1),
    Shrub.nn=
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(shrub_points),1),
    Hardwood.nn=
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(hardwood_points),1),
    Facilities.nn=
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(fire_suppression_facilities),3),
    WUI.nn=
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(wui_points),1),
    Fire.nn=
      nn_function(st_coordinates(st_centroid(fishnet_clipped)), st_coordinates(fire1013_points),5))

```


## Exploratory Analysis
- bar plots
The bar plots below visualize the continuous variables that were considered for the logistic model. The y-axis shows the mean value of each variable for eligible homeowners that do take the credit and those that do not take the credit. Variables that show significant differences between homeowners that do and do not take the credit will likely be useful for the model.

The bar plots illustrate the following:
- The mean number of.... 

- correlation matrix
- maybe some maps of categorical variables?

```{r exploratory, message=FALSE, warning=FALSE}
# DATA VISUALIZATIONS
##continuous variables
### Need to fix legend here
fishnet_clipped$FIRE <- ifelse(fishnet_clipped$Fire1418==1,"Fire","No_Fire")

fishnet_clipped %>% st_drop_geometry() %>%
  dplyr::select(FIRE, ELEVATION_AV, SLOPE_MEAN,MeanMaxTemp, MeanTemp1418,MeanHumidity1418,
                MeanPrecip1418,MeanWindSpeed1418, n_fires_intersections, Conifer.nn, Shrub.nn, Hardwood.nn, 
                Facilities.nn, WUI.nn) %>%
  rename("Elevation" = ELEVATION_AV, "Slope" = SLOPE_MEAN) %>%
  gather(Variable, value, -FIRE) %>%
  ggplot(aes(FIRE, value, fill=FIRE)) + 
  geom_bar(position = "dodge", stat = "summary", fun = "mean") + 
  facet_wrap(~Variable, scales = "free") +
  scale_fill_manual(values = palette2) +
  labs(x="Fire", y="Mean", 
       title = "Feature associations with the likelihood of Wildfire",
       subtitle = "(continous outcomes)") +
  theme(legend.position = "none")
```


```{r colinear, message=FALSE, warning=FALSE}
# Identifying Colinearity 
numericVars <- select_if(fishnet_clipped, is.numeric) %>% na.omit() %>% st_drop_geometry() %>%
  dplyr::select(ELEVATION_AV,SLOPE_MEAN,MeanTemp1418,MeanHumidity1418,
                MeanPrecip1418,MeanWindSpeed1418,MeanMaxTemp, n_fires_intersections,Conifer.nn, Shrub.nn, Hardwood.nn, 
                Facilities.nn, WUI.nn)

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
  labs(title = "Correlation across Characteristics") 
```


```{r exploratory_maps, message=FALSE, warning=FALSE}
Fire1418DF <- fishnet_clipped %>% dplyr::select(Fire1418, geometry)

ggplot() +
   geom_sf(data=Fire1418DF, aes(fill="transparent", color="black")) +
  geom_sf(data=fishnet_clipped, aes(fill=WUI_MAJ), color=NA) +
   scale_fill_discrete(name="WUI Type", labels=c("Missing Data","Not WUI","Influence Zone","Intermix","Interface")) +
  labs(title="Wildlife Urban Interface and Fire") +
  mapTheme()
  

```

## Describe the Spatial Process/ Space Time Process

# Model
- explain rare event problem
- list features in model: 
- FVEG_MAJ (vegetation type that makes up the majority of the grid cell)
- ELEVATION_AV (average elevation in the grid cell)
- SLOPE_MEAN (average slope in the grid cell)
- COVER_MAJ (majority land cover type in grid cell)
- COUNTY_ NAME 
- MeanTemp1418 (mean temperature during the fire season from 2014-2018)
- MeanHumidity1418 (mean humidity during the fire season from 2014-2018)
- MeanWindSpeed1418 (mean wind speed during the fire season from 2014-2018)
- n_fires_intersections (number of fires occurring in that grid cell)
- Conifer.nn (distance to nearest grid cell that is majority conifer)
- Hardwood.nn (distance to nearest grid cell that is majority hardwood)
- Facilities.nn (distance to 5 nearest fire suppression facilities)

- discuss statistical significance of variables

```{r model, message=FALSE, warning=FALSE}
# LOGISTIC MODEL
set.seed(1214)
trainIndex <- createDataPartition(fishnet_clipped$Fire1418, p = .65, 
                                 y = paste(fishnet_clipped$COVER_MAJ),
                                  list = FALSE,
                                  times = 1)

fireTrain <- fishnet_clipped[ trainIndex,] %>% st_drop_geometry()
fireTest  <- fishnet_clipped[-trainIndex,] %>% st_drop_geometry()

# MODEL
fireModel <- glm(Fire1418 ~ .,
                    data=fireTrain %>% 
                      dplyr::select(-ID,-Fire19,-weather_ID, -WUI_MAJ, -WUI.nn, 
                                    -MeanMaxTemp,-FIRE, -CoverCat, -ElevationBi,
                                    -Fire.nn, -SlopeCat, -MeanPrecip1418, -Max_Temp19, 
                                    -Mean_Temp19, -Mean_Precipitation19, 
                                    -Mean_Humidity19, -Mean_Wind_Speed19, -Fire1013, -Shrub.nn),
                    family="binomial" (link="logit"))

summary(fireModel)

## Adding Coefficients
x <- fireModel$coefficients
exp(x)
```


## Goodness of Fit
-McFadden: Higher McFadden pseudo r squared indicates better accuracy.Compared McFadden for different models to help choose
-Density of probabilities plot: Shows distribution of predicted probabilities for fire and no fire. Fire probabilities do not cluster around 1 because this is a rare event, very few grid cells have a high probability of wildfire. This is also shown in the histograms
-selecting a threshold: given the distribution ofpredicted probabilities for grid cells that have fires, the default threshold of .5 does not make sense. For our use case, we would prefer to over predict fires than to underpredict. 
-Confusion Matrix Finally, a confusion matrix for each model shows the number of true positives, true negatives, false positives, and false negatives for the model. A true positive here indicates that at a X% threshold, the model correctly predicted that a wildfire would occur in a given .5 x .5 mile grid cell. A true negative means that our model correctly predicted that a wildfire would not occur. False positive means the model predicted fire, but no fire occured. False negative indicates that the model predicted no fire, but there was fire. The confusion matrix output also shows the sensitivity, or true positive rate, and specificity, or true negative rate, for the model.

-ROC Curve: Thus far, we have used a threshold of 50% probability to determine whether a homeowner is predicted to take the credit. However, we may want to change that probability threshold. The receiver operating characteristic (ROC) curve visualizes trade-offs for different thresholds for the feature engineered model. As true positives in the model increase, the number of false positives also increase. That means that although HCD may help more people increase the value of their homes, they will waste more money marketing to people who will not end up taking the credit.

An area under the curve (AUC) of .5 would be a coin toss, meaning the true positive rate and false positive rate are equal. An AUC of 100% would be a perfect fit with 100% true negatives and 0 false positives. Here, we have an AUC between .5 and 1, which indicates that the model has a decent goodness of fit.
-discuss accuracy

```{r fit, message=FALSE, warning=FALSE}
## Fit metrics
pR2(fireModel)

## Prediction
testProbs <- data.frame(Outcome = as.factor(fireTest$Fire1418),
                        Probs = predict(fireModel, fireTest, type= "response"))


#Here we want more of a hump in the bottom plot around 1 to indicate that the reg is predictive
ggplot(testProbs, aes(x = Probs, fill = as.factor(Outcome))) + 
  geom_density() +
  facet_grid(Outcome ~ .) +
  scale_fill_manual(values = palette2) +
  labs(x = "Fire", y = "Density of probabilities",
       title = "Distribution of predicted probabilities by observed outcome",
       subtitle = "First Model") +
  theme(strip.text.x = element_text(size = 18),
        legend.position = "none")

# Histograms
testProbsFire <- testProbs %>% filter (Outcome=="1")
testProbsNoFire <- testProbs %>% filter (Outcome=="0")

hist(testProbsFire$Probs, 
     col="#F96167",
       main="Predicted Probabilities for Grid Cells with Fire",
     xlab="Probability")

hist(testProbsNoFire$Probs, 
     col="#FCE77D",
       main="Predicted Probabilities for Grid Cells with No Fire",
     xlab="Probability")

## Confusion matrix
### Might want to change this threshold, here a probability >50% if being predicted as takes credit
testProbs <- 
  testProbs %>%
  mutate(predOutcome  = as.factor(ifelse(testProbs$Probs > 0.2 , 1, 0)))

caret::confusionMatrix(testProbs$predOutcome, testProbs$Outcome, 
                       positive = "1")

# ROC Curve
## This us a goodness of fit measure, 1 would be a perfect fit, .5 is a coin toss
auc(testProbs$Outcome, testProbs$Probs)

ggplot(testProbs, aes(d = as.numeric(testProbs$Outcome), m = Probs)) +
  geom_roc(n.cuts = 50, labels = FALSE, colour = "#FE9900") +
  style_roc(theme = theme_grey) +
  geom_abline(slope = 1, intercept = 0, size = 1.5, color = 'grey') +
  labs(title = "ROC Curve - Model with Feature Engineering")
```

```{r 2019, message=FALSE, warning=FA:SE}
fireModel19 <- glm(Fire19 ~ .,
                 data=fishnet_clipped %>% st_drop_geometry()%>%
                   dplyr::select(-ID,-weather_ID, -Fire1418, -WUI_MAJ, -WUI.nn, 
                                 -MeanMaxTemp,-FIRE, -CoverCat, -ElevationBi,
                                 -Fire.nn, -SlopeCat, -MeanPrecip1418, -MeanMaxTemp, 
                                 -MeanTemp1418, -Mean_Precipitation19, 
                                 -MeanHumidity1418, -MeanWindSpeed1418, -Fire1013, -Shrub.nn),
                 family="binomial" (link="logit"))

summary(fireModel19)

testProbs19 <- data.frame(Outcome = as.factor(fishnet_clipped$Fire19),
                        Probs = predict(fireModel19, fishnet_clipped, type= "response"))

testProbs19nofire <- testProbs19 %>% filter(Outcome==0)
testProbs19fire <- testProbs19 %>% filter (Outcome==1)

hist(testProbs19nofire$Probs)
hist(testProbs19fire$Probs)

##0.006063904
mean(testProbs19nofire$Probs)

#0.2477918
mean(testProbs19fire$Probs)

## Confusion Matrix for 2019 at .25 threshold
testProbs19 <- 
  testProbs19 %>%
  mutate(predOutcome  = as.factor(ifelse(testProbs19$Probs > 0.25 , 1, 0)))

caret::confusionMatrix(testProbs19$predOutcome, testProbs$Outcome, 
                       positive = "1")
```

## Selecting a Threshold
```{r threshold, warning=FALSE, message=FALSE}
iterateThresholds <- function(data, observedClass, predictedProbs, group) {
#This function takes as its inputs, a data frame with an observed binomial class (1 or 0); a vector of predicted probabilities; and optionally a group indicator like race. It returns accuracy plus counts and rates of confusion matrix outcomes. It's a bit verbose because of the if (missing(group)). I don't know another way to make an optional parameter.
  observedClass <- enquo(observedClass)
  predictedProbs <- enquo(predictedProbs)
  group <- enquo(group)
  x = .01
  all_prediction <- data.frame()
  
  if (missing(group)) {
  
    while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x,2))
    
    all_prediction <- rbind(all_prediction,this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
  else if (!missing(group)) { 
   while (x <= 1) {
    this_prediction <- data.frame()
    
    this_prediction <-
      data %>%
      mutate(predclass = ifelse(!!predictedProbs > x, 1,0)) %>%
      group_by(!!group) %>%
      count(predclass, !!observedClass) %>%
      summarize(Count_TN = sum(n[predclass==0 & !!observedClass==0]),
                Count_TP = sum(n[predclass==1 & !!observedClass==1]),
                Count_FN = sum(n[predclass==0 & !!observedClass==1]),
                Count_FP = sum(n[predclass==1 & !!observedClass==0]),
                Rate_TP = Count_TP / (Count_TP + Count_FN),
                Rate_FP = Count_FP / (Count_FP + Count_TN),
                Rate_FN = Count_FN / (Count_FN + Count_TP),
                Rate_TN = Count_TN / (Count_TN + Count_FP),
                Accuracy = (Count_TP + Count_TN) / 
                           (Count_TP + Count_TN + Count_FN + Count_FP)) %>%
      mutate(Threshold = round(x,2))
    
    all_prediction <- rbind(all_prediction,this_prediction)
    x <- x + .01
  }
  return(all_prediction)
  }
}

whichThreshold <- 
  iterateThresholds(
     data=testProbs, observedClass = Outcome, predictedProbs = Probs)

whichThreshold <- 
  whichThreshold %>%
    dplyr::select(starts_with("Rate"), Threshold) %>%
    gather(Variable, Rate, Rate_TN,Rate_TP,Rate_FN,Rate_FP) 

ggplot(data=whichThreshold, aes(Threshold,Rate,color=Variable)) +
  geom_point()
  
whichThreshold %>%
  ggplot(.,aes(Threshold,Count,colour = Variable)) +
  geom_point() +
  #scale_colour_manual(values = palette5[c(5, 1:3)]) +    
  labs(title = "Revenue by confusion matrix type and threshold",
       y = "Count") +
  plotTheme() +
  guides(colour=guide_legend(title = "Confusion Matrix")) 

```

## Cross Validation
- discuss generalizability
The first outputs indicate the area under the Receiver Operating Characteristic (ROC) curve, the sensitivity, and the specificity for both models. Again we see that both models have high specificity (true negative rates), but low sensitivity. The second output visualizes these metrics for both models across folds. If the models were generalizable, we would expect all of these plots to be clustered around the mean. 
```{r cv, message=FALSE, warning=FALSE}
# K Fold Model Validation
ctrl <- trainControl(method = "cv", number = 100, classProbs=TRUE, summaryFunction=twoClassSummary)

cvFit <- train(Fire1418 ~ ., data = fishnet_clipped %>% st_drop_geometry() %>%
                 dplyr::select(
                   -ID,-Fire19,-weather_ID, -WUI_MAJ,
                   -MeanMaxTemp,-FIRE, -CoverCat, -ElevationBi,
                   -Fire.nn, -SlopeCat, -MeanPrecip1418, -Max_Temp19, 
                   -Mean_Temp19, -Mean_Precipitation19, 
                   -Mean_Humidity19, -Mean_Wind_Speed19, -Fire1013)%>%
                 dplyr::mutate(Fire1418=ifelse(Fire1418==1,"Fire","No_Fire")),
               method="glm", family="binomial",
               metric="ROC", trControl = ctrl)

cvFit

dplyr::select(cvFit$resample, -Resample) %>%
  gather(metric, value) %>%
  left_join(gather(cvFit$results[2:4], metric, mean)) %>%
  ggplot(aes(value)) + 
  geom_histogram(bins=35, fill = "#FF006A") +
  facet_wrap(~metric) +
  geom_vline(aes(xintercept = mean), colour = "#981FAC", linetype = 3, size = 1.5) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x="Goodness of Fit", y="Count", title="CV Goodness of Fit Metrics",
       subtitle = "Across-fold mean reprented as dotted lines") +
  plotTheme()
```

### Spatial Cross Validation
-discuss generalizability
```{r spatial_cv, message=FALSE, warning=FALSE}
# Model Validation
crossValidate <- function(dataset, id, dependentVariable, indVariables) {
  
  allPredictions <- data.frame()
  cvID_list <- unique(dataset[[id]])
  
  for (i in cvID_list) {
    
    thisFold <- i
    cat("This hold out fold is", thisFold, "\n")
    
    fold.train <- filter(dataset, dataset[[id]] != thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    fold.test  <- filter(dataset, dataset[[id]] == thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    
    regression <-
      glm(Fire1418 ~ ., family = "binomial", 
          data = fold.train %>% 
            dplyr::select(-geometry, -id))
    
    thisPrediction <- 
      mutate(fold.test, Prediction = predict(regression, fold.test, type = "response"))
    
    allPredictions <-
      rbind(allPredictions, thisPrediction)
    
  }
  return(st_sf(allPredictions))
}

reg.vars <- c("FVEG_MAJ","ELEVATION_AV","SLOPE_MEAN", "COVER_MAJ","COUNTY_NAME",
              "MeanTemp1418","MeanHumidity1418","MeanWindSpeed1418",
              "n_fires_intersections","Conifer.nn","Shrub.nn","Hardwood.nn",
              "Facilities.nn","WUI.nn")

reg.spatialCV <- crossValidate(
  dataset = fishnet_clipped,
  id = "COUNTY_NAME",
  dependentVariable = "Fire1418",
  indVariables = reg.vars) 

reg.spatialcv <-
  reg.spatialCV %>%
  dplyr::select(cvID = COUNTY_NAME, Fire1418, Prediction, geometry)

ggplot() +
  geom_sf(data = reg.spatialcv, aes(fill = Prediction), color = "transparent")+
  geom_sf(data = fire_perimeter1418, fill = "transparent", color = "red")

reg.spatialcv <-
  reg.spatialcv %>%
  mutate(Error = Prediction - Fire1418)

ggplot() +
  geom_sf(data = reg.spatialcv, aes(fill = Error), color = "transparent")
```

# Discussion
-critique strengths and weaknesses of model
- Overpredicting fire
- Weather idea generalized over several months and years, climate change, model needs to be updated for climate change
- difficulty predicting future bc a rare event
- accuracy went down as sensitivty went up, model predicting almost 0 fires had best accuracy
- giving risk score not binary outcome, risk always exists, we are trying to quantify amount of risk

# Conclusion

```{r conclusion, warning=FALSE, message=FALSE}

# Map of Risk
## Need to add geometry! 

testProbs_risk <- testProbs %>% mutate(Risk_Cat=
                                         case_when(Probs <= .09 ~ "Low",
                                                   Probs>.09 & Probs < .2 ~ "Medium",
                                                   Probs >=.2 ~ "High" ))

ggplot() +
  geom_sf(data=testProbs_risk, aes(fill=Risk_Cat))+
  mapTheme()

```

